---
title: "p8105_hw6_qra2000"
author: "Quinn Anderson"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%")

theme_set(theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5)))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

set.seed(1)
```

## Problem 2

First, load and tidy the Central Park weather data from class.

```{r, results=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

For this problem, we focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data: r^2 and log(beta_1 * beta_2)

The code chunk below draws 5000 samples and produces estimates of the two quantities noted above. The 2.5% and 97.5% quantiles are identified for both r^2 and log(beta_1 * beta_2). Please note, some estimates are negative for beta_2_hat, so the NAs were removed in order to produce the confidence intervals.

```{r}
bootstrap_weather = 
  weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm (tmax ~ tmin + prcp, data = df)),
    tidy_results = map(models, broom::tidy),
    glance_results = map(models, broom::glance)
  ) |> 
  select(tidy_results, glance_results, .id) |> 
  unnest(tidy_results, glance_results) |> 
  select(.id, term, estimate, r.squared) |> 
  group_by(.id) |> 
  summarize(
    log_beta12 = log(estimate[2]) + log(estimate[3]),
    r_squared = r.squared) |> 
  distinct()
```

The plot below is a distribution of log(beta_1 * beta_2). This distribution is skewed to the right.
```{r}
bootstrap_weather |> 
  ggplot(aes(x = log_beta12)) + 
  geom_density()
```

The plot below is a distribution of r^2. This is an approximately normal distribution that is slightly skewed to the right. 

```{r}
bootstrap_weather |> 
  ggplot(aes(x = r_squared)) + 
  geom_density()
```

Below is a 95% condifence interval for r^2 Log(beta_1 * beta_2).

```{r}
tibble(
    lower_beta = bootstrap_weather |> pull(log_beta12) |> quantile(0.025, na.rm = TRUE),
    upper_beta = bootstrap_weather |> pull(log_beta12) |> quantile(0.975, na.rm = TRUE),
    lower_r2 = bootstrap_weather |> pull(r_squared) |> quantile(0.025, na.rm = TRUE),
    upper_r2 = bootstrap_weather |> pull(r_squared) |> quantile(0.975, na.rm = TRUE)) |> 
  knitr::kable(digits = 2)
```



## Problem 3

First, we load and clean the data for regression analysis and convert the `babysex`, `frace`, and `mrace` variables to factor variables.

```{r}
bwt_df = read.csv("./data/birthweight.csv") |> 
  mutate(
    babysex = as.factor(case_when(babysex == 1 ~ "male",
                                  babysex == 2 ~ "female")),
    frace = as.factor(case_when(frace == 1 ~ "White",
                                frace == 2 ~ "Black",
                                frace == 3 ~ "Asian",
                                frace == 4 ~ "Puerto Rican",
                                frace == 8 ~ "Other",
                                frace == 9 ~ "Unknown")),
    mrace = as.factor(case_when(mrace == 1 ~ "White",
                                mrace == 2 ~ "Black",
                                mrace == 3 ~ "Asian",
                                mrace == 4 ~ "Puerto Rican",
                                mrace == 8 ~ "Other")))
```

Next, we propose a regression model based on a hypothesized structure for the factors that underly birthweight. My variables of interest are: family monthly income (in hundreds, rounded), average number of cigarettes smoked per day during pregnancy, mother's weight gain during pregnancy (pounds), and mother's race. 

We fit the simple linear regression to the dataset, summarize model estimates, and plot the model residuals against fitted values. 

```{r}
fit_1 = lm(bwt ~ fincome + smoken + wtgain + mrace, data = bwt_df)

fit_1 |> broom::tidy() |> knitr::kable(digits = 3)

bwt_df |> 
  add_predictions(fit_1) |> 
  add_residuals(fit_1) |> 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm", color = "red")
```

The plot shows there is a symmetric distribution above and below the line y = 0. This indicates that this is a normal distribution, so we proceed with the linear regression analysis. 

Next, we compare this model to two others: one using length at birth and gestational age as predictors, and one using head circumference, length, sex, and all interactions (including the three-way interaction) between these. We will make this comparison in terms of the cross-validated prediction error.

```{r}
cv_df =
  crossv_mc(bwt_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df |> 
  mutate(
    fit_1 = map(train, \(df) lm(bwt ~ fincome * smoken * wtgain, data = df)),
    fit_2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    fit_3 = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))) |> 
  mutate(
    rmse_fit_1 = map2_dbl(fit_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_fit_2 = map2_dbl(fit_2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_fit_3 = map2_dbl(fit_3, test, \(mod, df) rmse(model = mod, data = df)))
```

Next, we will display the RMSE results.

```{r}
cv_df |> 
  summarize(
    fit_1_mean_error = mean(rmse_fit_1),
    fit_2_mean_error = mean(rmse_fit_2),
    fit_3_mean_error = mean(rmse_fit_3)
  ) |> 
  knitr::kable(digits = 3)
```

The plot below shows the RMSE results in a violin plot:

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse ") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

